Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population

### 2023-12-11 manuscript/paper.Rmd
update

---
title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
subtitle: "Short title goes here!"
author:
- Amos Okutse, Brown University
- Second Author, Another Uni
- Graduate Student, Graduate School
abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.

**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio. Nam consequat non ante eget aliquam. Nulla facilisis tincidunt elit. Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut. Interdum et malesuada fames ac ante ipsum primis in faucibus. Suspendisse potenti. Duis vitae nibh mauris. Duis nec sem sit amet ante dictum mattis. Suspendisse diam velit, maximus eget commodo at, faucibus et nisi. Ut a pellentesque eros, sit amet suscipit eros. Nunc tincidunt quis risus suscipit vestibulum. Quisque eu fringilla massa."
keywords: "pandoc, r markdown, knitr"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
#mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
#bibliography: refs.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "../SleepBP/data/scrip2.R")
```

@@

-   importance of reproducibility
-   As a case on point ...?
-   Underscores the importance of reproducibility within the frame of
    statistical analysis. We use the study by to ... The current study
    builds on the paper by .. and starts in the manner of a replication
    study and extends the analyses therein. We focus on statistical
    analyses of these data and compare models based on varied
    specification of functional forms of the covariates.

@@

| Variable | Name                    | Description                                                                                                                                  |
|-------------|---------------|---------------------------------------------|
| SEQN     | sequence number         | Respondent number                                                                                                                            |
| SDMVPSU  | psu                     | Masked variance unit pseudo-PSU variable for variance estimation                                                                             |
| WTINT2YR | weights                 | Full sample 2-year interview weights                                                                                                         |
| SDMVSTRA | strata                  | Masked variance unit pseudo-stratum variable for variance estimation                                                                         |
| RIAGENDR | gender                  | Respondent's number                                                                                                                          |
| RIDAGEYR | age (yrs)               | Respondent's age in years                                                                                                                    |
| DMDMARTL | marital status          | Marital status                                                                                                                               |
| INDFMIN2 | income category         | Total family income (reported as a range value in dollars)                                                                                   |
| RIDRETH3 | race                    | Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category                                                    |
| DMDHHSZA | children \<5            | Number of children aged 5 years or younger in the household                                                                                  |
| DMDEDUC2 | education level         | What is the highest grade or level of school {you have/SP has} completed or the highest degree {you have/s/he has} received?                 |
| DMDCITZN | citizenship status      | {Are you/Is SP} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv             |
| SLD012   | sleep                   | Number of hours usually sleep on weekdays or workdays                                                                                        |
| BMXBMI   | bmi                     | Body mass index                                                                                                                              |
| ALQ121   | alcohol use             | In the past 12 months, how often did you drink any type of alcoholic beverage?                                                               |
| LBDHDDSI | hdl                     | High density lipoprotein                                                                                                                     |
| LBDSALSI | albumin                 | Albumin (g/L)                                                                                                                                |
| DIQ010   | diabetes                | Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?                                         |
| BPQ020   | hypertension            | {Have you/Has SP} ever been told by a doctor or other health professional that {you/s/he} had hypertension, also called high blood pressure? |
| URXCRS   | creatinine              | Creatinine, urine (umol/L)                                                                                                                   |
| SLQ040   | snort                   | In the past 12 months, how often did {you/SP} snort, gasp, or stop breathing while {you were/s/he was} asleep?                               |
| LBDTCSI  | total cholestrol levels | Total cholesterol (mmol/L)                                                                                                                   |
| LBXHGB   | hemoglobin              | Hemoglobin (g/dL)                                                                                                                            |
| LBXSASSI | AST                     | Aspartate aminotransferase                                                                                                                   |
| SMQ040   | smoke                   | Do you now smoke cigarettes?                                                                                                                 |
| BPXDI    | blood pressure          | Systolic and diastolic blood pressure taken as the average of the first three measurements                                                   |

```{r tabone, echo=FALSE}
#tab.one <- as_kable_extra(vars, caption = "Variable names and descriptions", 
#                 format = "latex", booktabs = TRUE) %>% 
#  kable_styling(latex_options = c("HOLD_position") %>% 
#  row_spec(c(0), bold = TRUE) %>% 
#  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
#  column_spec(3, width = "8cm")
#tab.one

#vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")

@@

With the purpose of identifying influential variables impacting blood
pressure, a variable selection process was conducted. The initial
approach involved employing best subset selection with 10-fold
cross-validation to ascertain the optimal number of subsets. Notably,
for systolic blood pressure, the forward best subset selection method
discerned 13 variables, whereas diastolic blood pressure exhibited 19
selected variables. However, a limitation inherent in forward best
subset selection is its inability to eliminate previously selected
features, potentially disregarding their relevance in light of newly
added variables. Consequently, this methodology may yield suboptimal
variable selections due to its lack of adaptability.

Table \@ref(tab:tabtwo) summarizes participant characteristics in the
2015 - 2018 NHANES survey cycles stratified by gender. Among the
participants, the proportion of males and females were 44.81% (n = 1378)
and 55.19% (n = 1509), respectively. With ethnicity, the proportion of
Mexican American, White and Black were 6.27%, 64.62% and 9.94%,
respectively. Overall, the mean (SD) values for age, albumin,
creatinine, SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 38.54 (17.72)
years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg,
68.47 (10.08) mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01)
IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2, respectively. Among the
participants, 66.01% were consumed alcohol drinkers, 2.61% were
diabetes, 3.16% were hypertension, 6.72% have snort or stop breathing.
17.99% were smokers. Sleep duration was divided into three groups, which
were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%,
53.21%, respectively.

Male and female participants differed significantly on characteristics
including

The univariate analysis of potential confounding factors is shown in
Table 2.

```{r tabtwo, echo=FALSE}
#tabsum %>% 
#  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
#  kable_styling(latex_options = c("HOLD_position"))

tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", 
                 longtable = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```

### Foward Stepwise Best Subset

```{=tex}
\begin{align*}
\text{sbp} = &\, 85.65 + 0.47 \times \text{BMI} + 1.49 \times \text{total\_chol} - 0.00 \times \text{creatinine} \\
&+ 3.06 \times \text{educ\_levelHighschool} + 2.00 \times \text{educ\_levelLess12grade} + 1.31 \times \text{educ\_levelsomCollege} \\
&+ 4.59 \times \text{genderMale} + 0.47 \times \text{age\_yr} + 2.05 \times \text{marital\_statusLiving with partner} \\
&- 0.12 \times \text{marital\_statusMarried} + 3.30 \times \text{marital\_statusNever married} \\
&- 0.99 \times \text{marital\_statusSeparated} + 7.01 \times \text{marital\_statusWidowed} - 1.30 \times \text{sleep>8hrs} \\
&- 2.82 \times \text{sleep6-8hrs} - 4.60 \times \text{raceMexican American} - 3.78 \times \text{raceOther} \\
&- 5.03 \times \text{raceWhite income\_categoryLower-middle income} - 3.02 \times \text{income\_categoryMiddle income} \\
&- 2.88 \times \text{income\_categoryUnknown/Refused} - 2.70 \times \text{income\_categoryUpper-middle income} \\
&- 4.30 \times \text{income\_categoryVaried/High income}
\end{align*}
```
```{=tex}
\begin{align*}
\text{dbp} = &\, 53.55 + 0.22 \times \text{BMI} + 0.78 \times \text{total\_chol} + 0.85 \times \text{hemoglobin} + 0.02 \times \text{albumin} \\
&+ 0.03 \times \text{AST} - 0.89 \times \text{diabetesNo} - 3.37 \times \text{diabetesYes} - 1.06 \times \text{children..51} \\
&- 1.67 \times \text{children..52} - 1.83 \times \text{children..53 or more} - 1.90 \times \text{marital\_statusLiving with partner} \\
&+ 0.00 \times \text{marital\_statusMarried} - 2.78 \times \text{marital\_statusNever married} + 0.91 \times \text{marital\_statusSeparated} \\
&+ 1.72 \times \text{marital\_statusWidowed} - 2.40 \times \text{sleep>8hrs} - 2.37 \times \text{sleep6-8hrs} - 2.76 \times \text{raceMexican American} \\
&+ 0.14 \times \text{raceOther} - 1.98 \times \text{raceWhite} - 2.16 \times \text{snortNot recorded} \\
&- 0.57 \times \text{snortYes} - 2.33 \times \text{alcoholNo drinking} - 1.79 \times \text{alcoholNot recorded}
\end{align*}
```
### LASSO

```{=tex}
\begin{align*}
\text{sbp} = &\, 84.13 \times \text{(Intercept)} + 0.52 \times \text{bmi} + 0.79 \times \text{hdl} + 1.43 \times \text{total\_chol} + 0.04 \times \text{AST} \\
&- 0.17 \times \text{diabetesNo} + 1.88 \times \text{diabetesYes} - 14.02 \times \text{citizenship\_statusDon't Know} \\
&- 0.59 \times \text{citizenship\_statusNot a citizen of the US} - 12.06 \times \text{citizenship\_statusRefused} \\
&+ 3.09 \times \text{educ\_levelHighschool} + 2.19 \times \text{educ\_levelLess12grade} + 1.16 \times \text{educ\_levelsomeCollege} \\
&- 0.99 \times \text{children..51} - 1.17 \times \text{children..52} + 0.90 \times \text{children..53 or more} \\
&+ 4.42 \times \text{genderMale} + 0.44 \times \text{age\_yr} + 1.77 \times \text{marital\_statusLiving with partner} \\
&+ 0.01 \times \text{marital\_statusMarried} + 2.95 \times \text{marital\_statusNever married} - 1.05 \times \text{marital\_statusSeparated} \\
&+ 7.19 \times \text{marital\_statusWidowed} - 1.47 \times \text{sleep>8hrs} - 2.95 \times \text{sleep6-8hrs} \\
&- 4.11 \times \text{raceMexican American} - 3.43 \times \text{raceOther} - 5.14 \times \text{raceWhite} \\
&+ 1.75 \times \text{smokeNot Smoking} + 0.22 \times \text{smokeSmoking} - 2.74 \times \text{snortNot recorded} \\
&- 0.61 \times \text{snortYes} - 1.69 \times \text{alcoholNo drinking} - 0.28 \times \text{alcoholNot recorded} \\
&- 1.62 \times \text{income\_categoryLower-middle income} - 3.04 \times \text{income\_categoryMiddle income} \\
&- 2.64 \times \text{income\_categoryUnknown/Refused} - 2.71 \times \text{income\_categoryUpper-middle income} \\
&- 4.54 \times \text{income\_categoryVaried/High income}
\end{align*}
```
```{=tex}
\begin{align*}
\text{dbp} = &\, 47.99 + 0.24 \times \text{bmi} + 0.80 \times \text{total\_chol} + 0.58 \times \text{hemoglobin} + 0.17 \times \text{albumin} \\
&+ 0.04 \times \text{AST} - 0.42 \times \text{diabetesNo} - 3.18 \times \text{diabetesYes} + 1.36 \times \text{citizenship\_statusDon't Know} \\
&- 0.60 \times \text{citizenship\_statusNot a citizen of the US} - 1.78 \times \text{citizenship\_statusRefused} \\
&- 0.11 \times \text{educ\_levelHighschool} - 0.92 \times \text{educ\_levelLess12grade} - 0.39 \times \text{educ\_levelsomeCollege} \\
&- 0.99 \times \text{children..51} - 1.39 \times \text{children..52} - 2.08 \times \text{children..53 or more} \\
&+ 0.59 \times \text{genderMale} + 0.01 \times \text{age\_yr} - 1.77 \times \text{marital\_statusLiving with partner} \\
&+ 0.12 \times \text{marital\_statusMarried} - 2.67 \times \text{marital\_statusNever married} + 0.94 \times \text{marital\_statusSeparated} \\
&+ 1.69 \times \text{marital\_statusWidowed} + 2.21 \times \text{cycle} - 2.21 \times \text{sleep>8hrs} \\
&- 2.25 \times \text{sleep6-8hrs} - 2.08 \times \text{raceMexican American} + 0.31 \times \text{raceOther} \\
&- 2.09 \times \text{raceWhite} + 0.44 \times \text{smokeNot Smoking} + 0.75 \times \text{smokeSmoking} \\
&- 2.15 \times \text{snortNot recorded} - 0.50 \times \text{snortYes} - 1.28 \times \text{alcoholNo drinking} \\
&- 1.17 \times \text{alcoholNot recorded}
\end{align*}
```
# Discussion and conclusion

>>>>

---
title: "Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
#subtitle: "Short title goes here!"
author:
- Amos Okutse, Brown University
- Monic Colon-Vargas, Brown University
abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.

**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio. Nam consequat non ante eget aliquam. Nulla facilisis tincidunt elit. Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut. Interdum et malesuada fames ac ante ipsum primis in faucibus. Suspendisse potenti. Duis vitae nibh mauris. Duis nec sem sit amet ante dictum mattis. Suspendisse diam velit, maximus eget commodo at, faucibus et nisi. Ut a pellentesque eros, sit amet suscipit eros. Nunc tincidunt quis risus suscipit vestibulum. Quisque eu fringilla massa."
keywords: "pandoc, r markdown, knitr"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
#mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
bibliography: bibfile.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
# source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

@@

The clinical importance of systolic blood pressure (SBP) stands as a
cornerstone extensively acknowledged and documented across the continuum
of medical literature's data collection [@gurven2012does; @su2022association]. Elevated SBP stands as a
precursor linked with a spectrum of critical health conditions, 
encompassing cardiovascular diseases, stroke, and renal impairment, among
others. This nexus between heightened SBP and these significant health
adversities underscores the gravity of SBP as a predictive marker for
adverse health outcomes. Noteworthy is the pivotal concern surrounding
the trajectory of SBP in the natural aging process, a trajectory
observed ubiquitously across diverse populations.

In western societies aged 40 years and above, a demonstrable pattern
emerges, revealing an approximate elevation of 7 mmHg in SBP per decade among individuals above 40 years. 
This discernible and consistent increment in SBP with
advancing age accentuates its profound impact within the broader context
of aging-related health dynamics. Similarly, while diastolic blood pressure
(DBP), demonstrates a concurrent rise correlating with age,
presents a substantially lower rate of increase when juxtaposed against
the ascending trajectory observed in SBP [@gurven2012does; @su2022association]. 
Thus, in light of the
significance attached to SBP dynamics within aging populations and the
intricate interplay between age, sleep, cholestrol levels and blood pressure alterations, our
investigative focus is oriented towards the exploration of variables
that exert potential influence on systolic blood pressure (SBP). This
deliberate focus aims to elucidate the multifaceted nature of factors
contributing to SBP variability, enabling a more comprehensive
understanding of its determinants within the framework of health and
aging.

<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->
<!-- contributor to systolic blood pressure (SBP), with individuals -->
<!-- exhibiting higher BMI values demonstrating a propensity for elevated -->
<!-- SBP. Gender also holds a discernible association with SBP dynamics. -->
<!-- Empirical evidence in CITE demonstrates that women typically manifest -->
<!-- lower SBP than men until the onset of menopause, after which women -->
<!-- exhibit higher SBP levels compared to men. Additionally, heightened -->
<!-- levels of cholesterol are closely linked to vascular damage. The -->
<!-- accumulation of cholesterol deposits and calcium within arteries -->
<!-- precipitates arterial stiffening, necessitating augmented cardiac -->
<!-- efforts to circulate blood, potentially culminating in elevated blood -->
<!-- pressure. This study endeavors to examine the interrelationships between -->
<!-- age and cholesterol levels, emphasizing consideration of potential -->
<!-- confounding variables. Table 1 comprehensively delineates the array of -->
<!-- variables under scrutiny, ensuring a comprehensive analysis of their -->
<!-- collective impact -->

@@

\begin{table}[H]

\caption{\label{tab:tabone}Variable descriptions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
\midrule
\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
\addlinespace
\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
\addlinespace
\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
\addlinespace
\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
\addlinespace
\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
\addlinespace
\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
\bottomrule
\end{tabular}
\end{table}

```{r tabone, echo=FALSE, eval=FALSE}
tab.one <- vars %>% 
  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
  kable_styling(latex_options = c("HOLD_position")) %>%
  row_spec(c(0), bold = TRUE) %>%
  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
  column_spec(3, width = "8cm")
tab.one
```

With the purpose of identifying influential variables impacting blood
pressure, a variable selection process was conducted. The initial
approach involved employing best subset selection with 10-fold
cross-validation to ascertain the optimal number of subsets. Notably,
for systolic blood pressure, the forward best subset selection method
discerned 13 variables, whereas diastolic blood pressure exhibited 19
selected variables. However, a limitation inherent in forward best
subset selection is its inability to eliminate previously selected
features, potentially disregarding their relevance in light of newly
added variables. Consequently, this methodology may yield sub-optimal
variable selections due to its lack of adaptability.

## Model Evaluations

Table \@ref(tab:tabmodel) summarizes the performance metrics across the different models fitted to the data in our investigation.

\begin{table}

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}

```{r tabmodel, eval=FALSE, echo=FALSE}
## Model performance
tabmodel <- model.checks %>% 
  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
tabmodel
```

## Descriptive statistics

Table \@ref(tab:tabtwo) summarizes participant characteristics in the
2015 - 2018 NHANES survey cycles stratified by gender. The 2015-2016 survey cycle comprised 53.77% of the total analyzed sample whereas the 2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample. 
Among the
participants, the proportion of males and females were 53.76% (n = 1063)
and 46.23% (n = 914), respectively. With race, the proportion of
Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
respectively. Overall, the mean (SD) values for age, albumin, 
SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43)
years, 43.04 (3.59) g/L, 118.61 (16.41) mmHg,
69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81 (1.02) mmol/L, 23.47 (13.17)
IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03) kg/m2, respectively. 
Among the
participants, 73.70% were alcohol drinkers, 5.06% were
diabetic, 7.79% were hypertensive, 13.76% experienced snorting whereas
21.60% were smokers. Sleep duration was divided into three groups, which
were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 7.89%, 40.06%,
52.05%, respectively. Male and female participants differed significantly on all characteristics
except citizenship status, and the cycle of the survey.

```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase. SBP = Systolic blood pressure, DBP = Diastolic blood pressure.", 
                 longtable = TRUE)
```

## Univariate regression analysis

Table \@ref(tab:tabthree) presents the summary of univariate regression analysis results exploring the association between the exposures, selected potential confounding variables and systolic blood pressure. We found substantial associations between ...

```{r tabthree, echo=FALSE}
## univariate analysis
sbp.univariate %>% 
  as_kable_extra(format = "latex", booktabs = TRUE, digits = 2, caption = "Univariate analyses of exposures, potential confounding factors and systolic blood pressure. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase.", longtable = TRUE)
```

## Multivariable Regression Analysis

```{r tabfour, echo=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

## Subgroup analysis by gender

# Discussion and conclusion
